### 2022-09-04 appendix_kronecker/appendix.tex
is it finished?? :o

%!TEX root = ../thesis.tex
\chapter{Kronecker properties}\label{appendix: kronecker}
Here we detail some properties of Kronecker sums, products, and exponential (see \citep{MEinAP}, Appendix A.4).

Let 
\[\bs A = \left[\begin{array}{ccc}a_{11} & \dots & a_{1m}\\\hdots & & \hdots \\ a_{n1}&\dots & a_{nm}\end{array}\right]
\qquad
\bs{B} = \left[\begin{array}{ccc}b_{11} & \dots & b_{1m'}\\\hdots & & \hdots \\ b_{n'1}&\dots & b_{n'm'}\end{array}\right]\]
be matrices with dimensions \(n\times m\) and \(n'\times m'\), respectively. The operator \(\otimes\) is the Kronecker product of two matrices; 
\[\bs A\otimes \bs{B} = \left[\begin{array}{ccc}a_{11}\bs{B} & \dots & a_{1m}\bs{B}\\\hdots & & \hdots \\ a_{n1}\bs{B}&\dots & a_{nm}\bs{B}\end{array}\right],\]
which is an \(nn'\times mm'\) matrix.

Let \(\bs{C},\bs{D}\)  be matrices with dimensions \(m\times k\) and \(m'\times k'\). A property of the Kronecker Product is 
\begin{align}
	\left(\bs A\otimes \bs{B}\right)\left(\bs{C}\otimes \bs{D}\right) &= \bs A\bs{C}\otimes \bs{B}\bs{D}.\label{eqn:mpr}\tag{Mixed Product Rule}
\end{align}
\begin{proof}
	The proof follows from 
	\begin{align*}
		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j}\\\vdots\\c_{nj}\bs{D} \end{array}\right] 
		%
		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
	\end{align*}
\end{proof}

If \(\bs A\) and \(\bs{B}\) are invertible matrices, then 
\begin{align}\label{eqn:kron inverse}
	\left(\bs A\otimes \bs{B}\right)^{-1} = \bs A^{-1}\otimes \bs{B}^{-1}.
\end{align}

Let \(\bs A\) and \(\bs{B}\) be \(n\times n\) and \(m\times m\) matrices, respectively. The Kronecker sum of \(\bs A\) and \(\bs{B}\) is denoted by \(\oplus\) and defined as 
\[\bs A\oplus \bs{B} := \bs A\otimes \bs{I}_{m} + \bs{I}_{n}\otimes \bs{B}.\]

The exponential of a square matrix \(\bs B\) is \[e^{\bs B} := \sum_{n=0}^\infty \cfrac{1}{n!}\bs B^n.\]

A property of the Kronecker sum is 
\begin{align}\label{eqn:kron exp}
	e^{\bs A\oplus \bs{B}}= e^{\bs A}\otimes e^{\bs{B}}.
\end{align}
\begin{proof}
	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\). The latter follows from the fact that \(\bs{I}_n\otimes \bs{B}\) is a block diagonal matrix with blocks \(\bs{B}\), hence its exponential is also block diagonal with blocks equal to the exponential of \(\bs{B}\). The former follows from 
	\begin{align}
	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
	\end{align}

	Therefore 
	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
	and the result follows by the mixed product rule. 
\end{proof}

\begin{lem}\label{lem: lst mpr}
	Let \(\bs{T}\) and \(\bs{C}\) be \(n\times n\), square matrices with \(\bs{C}\) diagonal and invertible; let \(\bs{S}\) be a \(p\times p\) matrix. Further, suppose \(\left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]\) is invertible for \(\lambda>0\). Then
\begin{align}
	&\int_{t=0}^\infty e^{-\lambda t}  e^{{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t}} \wrt t 
	%
	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}.  \label{eqn:lstsimplify}\end{align}
\end{lem}
\begin{proof}
	Computing the integral on the left-hand side and then factorising the result and using the~\ref{eqn:mpr} multiple times gives
	\begin{align}
            	\int_{t=0}^\infty e^{-\lambda t} e^{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t} \wrt t\nonumber 
            	%
            	&= - \left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]^{-1}
		%
		\\&= -  \left[\bs{T}\otimes \bs{I} + \left(\bs{C}\otimes \bs{I}\right)\left(\bs{I}\otimes \bs{S}\right) - \lambda \bs{I}\right]^{-1}\nonumber
		%
		\\&= -  \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}\otimes \bs{I}\right)^{-1}\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}\otimes \bs{I}\right)^{-1}\lambda \bs{I}\right)\right]^{-1}. \label{eqn: ref this one 12}
		%
	\end{align}
	By Equation~(\ref{eqn:kron inverse}) and since \(\bs{C}\) is invertible, (\ref{eqn: ref this one 12}) is equal to
	\begin{align}
		& - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\otimes \bs{I}\right)\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\otimes \bs{I}\right)\lambda \bs{I}\right)\right]^{-1}. \label{eqn: ref this one 13}
		%
	\end{align}
	{Using the~\ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
	\begin{align}
		&- \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\bs{T}\right)\otimes \bs{I} + \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\lambda \bs{I}\right)\otimes \bs{I}\right)\right]^{-1} \nonumber
		%
		\\&= - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right)\right]^{-1} \nonumber
		%
		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1} \nonumber
		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1},\label{eqn: is an integral}
	\end{align}
	by definition of the Kronecker sum.

	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore, (\ref{eqn: is an integral}) is 
	\begin{align*}
		-\left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1}
		&= \int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x\right)\oplus \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1}.
	\end{align*}
	{Using the rule in Equation~(\ref{eqn:kron exp}) gives }
	\begin{align*}
		&\int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)x}\otimes e^{ \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1},
	\end{align*}
	which is the result.
\end{proof}

%\section{Matrix exponentials}\label{appendix: matrix exponentials}
%This appendix contains some technical results regarding the matrix exponential.

>>>>

%!TEX root = ../thesis.tex
\chapter{Algebraic manipulations of certain Kronecker products}\label{appendix: kronecker}
%\section{Matrix exponentials}\label{appendix: matrix exponentials}
%This appendix contains some technical results regarding the matrix exponential.

### 2022-09-04 chapter1/chapter.tex
is it finished?? :o

\section{Convergence theorems}
We use the following convergence theorems to help us prove that the QBD-RAP scheme converges weakly to the distribution of the fluid queue. The first result we state, the Portmanteau Theorem, is a sweeping statement about convergence of measures. First, let's define the notion of weak convergence. We follow \cite{portmanteaubook}.

Denote by \(\mathcal M_f(E)\) the set of all finite measures on \((E,\mathcal E)\), where \(E\) is a non-empty set and \(\mathcal E\) is a \(\sigma\)-algebra. Further, denote by \(C_b(E)\) the set of continuous bounded functions on \(E\). Let \(\mu,\,\mu_1,\,\mu_2,...\in \mathcal M_f(E)\). We say that \(\{\mu_n\}_{n\in\mathbb N}\) converges weakly to \(\mu\), formally \(\mu_n\to \mu\) weakly as \(n\to\infty\), if 
\[\int f \wrt \mu_n \to \int f\wrt \mu,\mbox{ for all }f \in C_b(E).\]
We use part of The Portmanteau Theorem in Chapter~\ref{ch: global results}. Let \(\mathcal M_{\leq 1}(E):=\{\mu \in \mathcal M_f(E)\mid \mu(E)\leq 1\},\) the set of all sub-probability measures on \((E,\mathcal E)\). 
\begin{thm}[[Part of] The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
	Let \(E\) be a metric space and let \(\mu,\, \mu_1,\,\mu_2,... \in \mathcal M_{\leq 1}(E)\). The following are equivalent.
	\begin{enumerate}
		\item[(i)] \(\mu_n\to \mu\) weakly as \(n\to \infty\). 
		\item[(ii)] \(\int f\wrt \mu_n \to \int f\wrt \mu\) for all bounded, Lipschitz continuous \(f\).
		\\ \(\vdots\) 
	\end{enumerate}
\end{thm}
There are 8 parts to The Portmanteau Theorem in \cite{portmanteaubook}, Theorem~13.16. Here we only quote to relevant parts; the other 6 parts require us to define other concepts which are not relevant to this thesis, so they are omitted. See \cite{portmanteaubook}, Theorem~13.16 for details.

@@

\subsection{Briefly, on accuracy}
As a rough guide on accuracy we now paraphrase some known results on the accuracy of the DG method from \cite[Sections~5.5,~5.8]{nodalDGBook} for a similar kind of problem to the one under study here. For certain scalar conservation-law problems, under certain regularity conditions on the continuity and bounded-ness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}. Applying this result to the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.

@@

>>>>

@@

\subsection{Briefly, on accuracy}
As a rough guide on accuracy we now paraphrase some known results on the accuracy of the DG method from \cite[Sections~5.5,~5.8]{nodalDGBook} for a similar kind of problem to the one under study here. For certain scalar conservation-law problems, under certain regularity conditions on the continuity and boundedness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}. Applying this result to the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.

@@

\section{Kronecker products and related results}
Here we describe Kronecker products and sums and some of their properties (see \cite{MEinAP}, Appendix A.4, for further details). We use the results in this section to manipulate certain matrix expressions in Chapter~\ref{sec: conv}.

Let 
\[\bs A = \left[\begin{array}{ccc}a_{11} & \dots & a_{1m}\\\hdots & & \hdots \\ a_{n1}&\dots & a_{nm}\end{array}\right]
\qquad
\bs{B} = \left[\begin{array}{ccc}b_{11} & \dots & b_{1m'}\\\hdots & & \hdots \\ b_{n'1}&\dots & b_{n'm'}\end{array}\right]\]
be matrices with dimensions \(n\times m\) and \(n'\times m'\), respectively. The operator \(\otimes\) is the Kronecker product of two matrices; 
\[\bs A\otimes \bs{B} = \left[\begin{array}{ccc}a_{11}\bs{B} & \dots & a_{1m}\bs{B}\\\hdots & & \hdots \\ a_{n1}\bs{B}&\dots & a_{nm}\bs{B}\end{array}\right],\]
which is an \(nn'\times mm'\) matrix.

Let \(\bs{C},\bs{D}\)  be matrices with dimensions \(m\times k\) and \(m'\times k'\). A property of the Kronecker Product is 
\begin{align}
	\left(\bs A\otimes \bs{B}\right)\left(\bs{C}\otimes \bs{D}\right) &= \bs A\bs{C}\otimes \bs{B}\bs{D}.\label{eqn:mpr}\tag{Mixed Product Rule}
\end{align}
\begin{proof}
	The proof follows from 
	\begin{align*}
		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j}\\\vdots\\c_{nj}\bs{D} \end{array}\right] 
		%
		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
	\end{align*}
\end{proof}

If \(\bs A\) and \(\bs{B}\) are invertible matrices, then 
\begin{align}\label{eqn:kron inverse}
	\left(\bs A\otimes \bs{B}\right)^{-1} = \bs A^{-1}\otimes \bs{B}^{-1}.
\end{align}

Let \(\bs A\) and \(\bs{B}\) be \(n\times n\) and \(m\times m\) matrices, respectively. The Kronecker sum of \(\bs A\) and \(\bs{B}\) is denoted by \(\oplus\) and defined as 
\[\bs A\oplus \bs{B} := \bs A\otimes \bs{I}_{m} + \bs{I}_{n}\otimes \bs{B}.\]

The exponential of a square matrix \(\bs B\) is \[e^{\bs B} := \sum_{n=0}^\infty \cfrac{1}{n!}\bs B^n.\]

A property of the Kronecker sum is 
\begin{align}\label{eqn:kron exp}
	e^{\bs A\oplus \bs{B}}= e^{\bs A}\otimes e^{\bs{B}}.
\end{align}
\begin{proof}
	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\). The latter follows from the fact that \(\bs{I}_n\otimes \bs{B}\) is a block diagonal matrix with blocks \(\bs{B}\), hence its exponential is also block diagonal with blocks equal to the exponential of \(\bs{B}\). The former follows from 
	\begin{align}
	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
	\end{align}

	Therefore 
	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
	and the result follows by the mixed product rule. 
\end{proof}

\begin{lem}\label{lem: lst mpr}
	Let \(\bs{T}\) and \(\bs{C}\) be \(n\times n\), square matrices with \(\bs{C}\) diagonal and invertible; let \(\bs{S}\) be a \(p\times p\) matrix. Further, suppose \(\left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]\) is invertible for \(\lambda>0\). Then
\begin{align}
	&\int_{t=0}^\infty e^{-\lambda t}  e^{{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t}} \wrt t 
	%
	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}.  \label{eqn:lstsimplify}\end{align}
\end{lem}
\begin{proof}
	Computing the integral on the left-hand side and then factorising the result and using the~\ref{eqn:mpr} multiple times gives
	\begin{align}
            	\int_{t=0}^\infty e^{-\lambda t} e^{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t} \wrt t\nonumber 
            	%
            	&= - \left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]^{-1}
		%
		\\&= -  \left[\bs{T}\otimes \bs{I} + \left(\bs{C}\otimes \bs{I}\right)\left(\bs{I}\otimes \bs{S}\right) - \lambda \bs{I}\right]^{-1}\nonumber
		%
		\\&= -  \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}\otimes \bs{I}\right)^{-1}\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}\otimes \bs{I}\right)^{-1}\lambda \bs{I}\right)\right]^{-1}. \label{eqn: ref this one 12}
		%
	\end{align}
	By Equation~(\ref{eqn:kron inverse}) and since \(\bs{C}\) is invertible, (\ref{eqn: ref this one 12}) is equal to
	\begin{align}
		& - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\otimes \bs{I}\right)\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\otimes \bs{I}\right)\lambda \bs{I}\right)\right]^{-1}. \label{eqn: ref this one 13}
		%
	\end{align}
	{Using the~\ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
	\begin{align}
		&- \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\bs{T}\right)\otimes \bs{I} + \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\lambda \bs{I}\right)\otimes \bs{I}\right)\right]^{-1} \nonumber
		%
		\\&= - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right)\right]^{-1} \nonumber
		%
		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1} \nonumber
		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1},\label{eqn: is an integral}
	\end{align}
	by definition of the Kronecker sum.

	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore, (\ref{eqn: is an integral}) is 
	\begin{align*}
		-\left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1}
		&= \int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x\right)\oplus \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1}.
	\end{align*}
	{Using the rule in Equation~(\ref{eqn:kron exp}) gives }
	\begin{align*}
		&\int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)x}\otimes e^{ \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1},
	\end{align*}
	which is the result.
\end{proof}

Laplace transforms of probability measures with non-negative support and where the Laplace transform variable, \(\lambda\), is real and non-negative can be given a stochastic interpretation. Let \(W\) be a random variable with distribution function \(F_W(w)= \mathbb P(W<w)\), then \(\displaystyle\int_{w=0}^\infty e^{-\lambda w} \wrt F_W(w) = \mathbb P(W < E^{\lambda})\). That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(W\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.

\section{Convergence theorems}
We use the following convergence theorems to help us prove that the QBD-RAP scheme converges weakly to the distribution of the fluid queue. The first result we state, the Portmanteau Theorem, is a sweeping statement about convergence of measures. First, let's define the notion of weak convergence. We follow \cite{billingsleyconvergence}.

Let \(S\) be a metric space and let \(\mathcal S\) be the Borel \(\sigma\)-algebra generated by the open sub-sets of \(S\). A probability measure, \(P\), is a function which maps elements of \(\mathcal S\) (elements of \(\mathcal S\) are sets) to real numbers in the interval \([0,1]\), with \(P(S)=1\). Further, \(P\) is \emph{countably-additive}, which means that for any countable collection of \emph{disjoint} sets \(A_1, A_2,...\in\mathcal S\), then 
\[P\left(\bigcup\limits_{n=1}^\infty A_n\right) = \sum\limits_{n=1}^\infty P\left(A_n\right).\]
Define the notation \(Pf = \int_S f \wrt P\) where \(f:S\to \mathbb R\) is a function. Let \(P_1,P_2,...\), be a sequence of probability measures. For a given function \(f\), the sequence \(P_1f,P_2f,...\), is simply a sequence of real numbers. The sequence of probability measures \(P_1,P_2,....\), is said to \emph{converge weakly} to \(P\) if \(P_nf\to Pf\) for every bounded continuous real function \(f\). \cite{billingsleyconvergence} uses the notation \(P_n\Rightarrow P\) to denote this weak convergence.

We need a few more definitions before stating The Portmanteau Theorem. A set \(A\in\mathcal S\) is said to be a \(P\)-\emph{continuity set} if \(\partial A\) (the boundary of the set \(A\)) satisfies \(P(\partial A)=0\). Define the \emph{limit inferior} and \emph{limit superior} of a sequence of real numbers \(x_1,x_2,...,\) as
\begin{align*}
	\liminf_{n\to\infty} x_n &:= \lim_{n\to \infty}	\left(\inf_{m\geq n}x_m\right)
	\intertext{and}
	\\\limsup_{n\to\infty} x_n &:= \lim_{n\to \infty}	\left(\sup_{m\geq n}x_m\right),
\end{align*}
respectively.

\begin{thm}[Portmanteau Theorem I, Theorem~2.1 of \cite{billingsleyconvergence}]
	These five conditions are equivalent.
	\begin{enumerate}
		\item[(i)] \(P_n\Rightarrow P\). 
		\item[(ii)] \(P_nf\to Pf\) for all bounded, uniformly continuous \(f\).
		\item[(iii)] \(\limsup_n P_n F \leq P F\) for all closed [sets] \(F\).
		\item[(iv)] \(\liminf_n P_n G \leq P G\) for all open [sets] \(G\).
		\item[(v)] \(P_n A \to P A\) for all \(P\)-continuity sets \(F\).
	\end{enumerate}
\end{thm}
Some authors, such as \cite{portmanteaubook}, also include the following equivalent condition in the Portmanteau Theorem. 
\begin{thm}[Portmanteau Theorem II, Theorem~13.16 of \cite{portmanteaubook}]\label{thm: Portmanteau}
	\(\,\)
	\begin{enumerate}
		\item[(vi)] \(P_nf \to P f\) for all bounded, Lipschitz continuous \(f\).
	\end{enumerate}
\end{thm}

Recall that a stochastic process is a sequence of random variables \(\{X(t)\}_{t\in\mathcal T}\). The \emph{finite-dimensional distributions} of \(\{X(t)\}_{t\in\mathcal T}\) are the joint distributions of the random vector \((X(t_1),X(t_2),...,X(t_n))\), where \(t_1,t_2,...,t_n\in \mathcal T\) is a finite collection of times.

Key concepts in establishing weak convergence of stochastic processes are the notions \emph{tight} and \emph{relatively compact}. Again, we follow \cite{billingsleyconvergence}. A probability measure \(P\) is said to be \emph{tight} if for each \(\varepsilon\) there exists a compact set \(K\) such that \(P(K)>1-\varepsilon\). A family of probability measures, \(\Pi\), is said to be tight if for every \(\varepsilon\) there exists a compact set \(K\) such that \(P(K)>1-\varepsilon\) for every probability measure \(P\) in \(\Pi\). The family \(\Pi\) is \emph{relatively compact} if every sequence of elements of \(\Pi\) contains a subsequence which converges weakly.

Let \(P_1,P_2,...,\) and \(P\) be probability measures of stochastic processes. \cite{billingsleyconvergence} provides the following result.
\begin{thm}\label{thm: aofaa}
	If \(\{P_n\}\) is relatively compact and the finite-dimensional distributions of \(P_n\) converge weakly to those of \(P\), then \(P_n\Rightarrow_n P\). 
\end{thm}
Further, the condition that \(\{P_n\}\) is relatively compact can be replaced by tightness due Prohorov's Theorem \citep{billingsleyconvergence}. 
\begin{thm}
	If \(\Pi\) is tight, then it is relatively compact.
\end{thm}

% Denote by \(\mathcal M_f(E)\) the set of all finite measures on \((E,\mathcal E)\), where \(E\) is a non-empty set and \(\mathcal E\) is a \(\sigma\)-algebra. Further, denote by \(C_b(E)\) the set of continuous bounded functions on \(E\). Let \(\mu,\,\mu_1,\,\mu_2,...\in \mathcal M_f(E)\). We say that \(\{\mu_n\}_{n\in\mathbb N}\) converges weakly to \(\mu\), formally \(\mu_n\to \mu\) weakly as \(n\to\infty\), if 
% \[\int f \wrt \mu_n \to \int f\wrt \mu,\mbox{ for all }f \in C_b(E).\]
% We use part of The Portmanteau Theorem in Chapter~\ref{ch: global results}. Let \(\mathcal M_{\leq 1}(E):=\{\mu \in \mathcal M_f(E)\mid \mu(E)\leq 1\},\) the set of all sub-probability measures on \((E,\mathcal E)\). 
% \begin{thm}[(Part of) The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
% 	Let \(E\) be a metric space and let \(\mu,\, \mu_1,\,\mu_2,... \in \mathcal M_{\leq 1}(E)\). The following are equivalent.
% 	\begin{enumerate}
% 		\item[(i)] \(\mu_n\to \mu\) weakly as \(n\to \infty\). 
% 		\item[(ii)] \(\int f\wrt \mu_n \to \int f\wrt \mu\) for all bounded, Lipschitz continuous \(f\).
% 		\\ \(\vdots\) 
% 	\end{enumerate}
% \end{thm}
% There are 8 parts to The Portmanteau Theorem in \cite{portmanteaubook}, Theorem~13.16. Here we only quote to relevant parts; the other 6 parts require us to define other concepts which are not relevant to this thesis, so they are omitted. See \cite{portmanteaubook}, Theorem~13.16 for details.
